### 2023-10-29 thesis/sections/5-grounding.tex
results

\begin{table}[ht]
    \centering
    \begin{tabular}{rcc}
        \toprule
                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
        \textbf{Dale-2}      & 50\%                       & 99\%              \\
        \textbf{Dale-5}      & 22\%                       & 94\%              \\
        \textbf{CLEVR color} & 12\%                       & 93\%              \\
        \bottomrule
    \end{tabular}
    \caption{Results of the bounding box classifier after 10 epochs}
    \label{tab:results_bounding_box_classifier}
\end{table}

@@

In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input.
The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e_i$ with a linear layer.
All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $d$ using another linear layer.
This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
Tokens used in the LSTM are embedded with embedding dimension $e_t$.
During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.

@@

In the second setup two different models are compared against each other.
The first model, the \emph{basic RE generator} acts as the baseline and only receives the complete image as the input.
The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
The resulting vector is then reduced to decoder output dimension $d$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
The LSTM is trained in the same way as above.

@@

For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
The following values for the variables are compared:
\begin{itemize}
    \item $e_i$ (only for bounding boxes): 100, 500, 1000
    \item $d$: 100, 500, 1000
    \item $e_t$: 10, 15, 30
\end{itemize}

@@

It could be the case that the model predicted the correct shape, but wrong color.
Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
% DK: TODO
The correct caption is \emph{cylinder}.
If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
To mitigate this, the accuracies for each class are included, as well as the macro average.
% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
This can give a better understanding of the errors the model makes.
The same is done for \emph{precision} and \emph{recall}.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{rccc}
        \toprule
                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
        \bottomrule
    \end{tabular}
    \caption{Results of the caption generator, based on ResNet-3}
    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1. It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
    % DK: might be the case. in the last language game i use only a linear layer instead of convolutions, but it doesn't help. This however might also be due to other problems. Time? (QUESTION, future work)
    % SD: I would present these for individual classes.
    % DK: I only have accuracy scores at. Time? (QUESTION)
    \label{tab:results_caption-generator}
\end{table}

>>>>

\begin{table}[ht]
    \centering
    \begin{tabular}{c|c|c|c}
        \toprule
               & \textbf{Dale-2}   & \textbf{Dale-5}   & \textbf{CLEVR Color} \\\cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}
        $e$    & \textbf{Accuracy} & \textbf{Accuracy} & \textbf{Accuracy}    \\\midrule
        {2}    & {92\%}            & {72\%}            & {40\%}               \\
        {10}   & \textbf{100\%}    & {94\%}            & {92\%}               \\
        {50}   & {99\%}            & \textbf{95\%}     & \textbf{94\%}        \\
        {100}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
        {500}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
        {1000} & \textbf{100\%}    & {94\%}            & \textbf{94\%}        \\
        \bottomrule
    \end{tabular}
    \caption{Accuracy scores of the object identifier: $e$ are different embedding sizes}
    \label{tab:results_bounding_box_classifier}
\end{table}

@@

In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input.
The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e$ with a linear layer.
All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $LSTM_o$ using another linear layer.
This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
Tokens used in the LSTM are embedded with embedding dimension $LSTM_e$.
During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.

@@

In the second setup two different models are compared against each other.
The first model, the \emph{basic RE generator} acts as the baseline and only receives the complete image as the input.
The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
The resulting vector is then reduced to decoder output dimension $LSTM_o$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
The LSTM is trained in the same way as above.

@@

For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
The following values for the variables are compared:
\begin{itemize}
    \item $e$ (only for bounding boxes): 100, 500, 1000
    \item $LSMT_o$: 100, 500, 1000
    \item $LSTM_e$: 10, 15, 30
\end{itemize}

@@

It could be the case that the model predicted the correct shape, but wrong color.
Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
% DK: TODO
The correct caption is \emph{cylinder}.
If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
To mitigate this, the accuracies for each class are included, as well as the macro average.
% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
This can give a better understanding of the errors the model makes.
The same is done for \textbf{precision} and \textbf{recall}.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{ccc|ccc|ccc|ccc}
        \toprule
               &          &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                 \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
        $e$    & $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}     & \textbf{NT}  & \textbf{Acc.} & \textbf{F1}   & \textbf{NT}  \\\midrule
        {100}  & {100}    & {10}     & {97\%}                              & {97,5\%}                            & {1\%}                                    & {65\%}        & {88\%}          & {2\%}        & {92\%}        & {81\%}        & {3\%}        \\
        {100}  & {100}    & {15}     & {97\%}                              & {98\%}                              & {0\%}                                    & {62\%}        & {86,47\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
        {100}  & {100}    & {30}     & {98\%}                              & {98\%}                              & {1\%}                                    & {65\%}        & {87,99\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
        {100}  & {500}    & {10}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {67\%}        & {88\%}          & {2\%}        & {93\%}        & {82\%}        & {4\%}        \\
        {100}  & {500}    & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
        {100}  & {1000}   & {10}     & {98\%}                              & {98\%}                              & {0\%}                                    & {68\%}        & {88\%}          & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
        {500}  & {500}    & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {67\%}        & {88,99\%}       & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
        {500}  & {1000}   & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
        {1000} & {100}    & {10}     & {97\%}                              & {97,99\%}                           & {0\%}                                    & {59\%}        & {85,5\%}        & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
        {1000} & {100}    & {30}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {61\%}        & {87\%}          & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
        {1000} & {1000}   & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & {93\%}        & {82\%}        & {4\%}        \\
        {1000} & {1000}   & {15}     & {99\%}                              & {98,5\%}                            & {0\%}                                    & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
        \bottomrule
    \end{tabular}
    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
    \label{tab:results_bounding-box-caption-generator}
\end{table}

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|ccc|ccc|ccc}
        \toprule
                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                         \\\cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
        $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1} & \textbf{NT} & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule
        {100}    & {10}     & {\%}                                & {\%}                                & {\%}                                     & {65\%}        & {88\%}      & {2\%}       & {92\%}        & {81\%}      & {3\%}       \\
        {100}    & {15}     & {78\%}                              & {71,99\%}                           & {11\%}                                   & {62\%}        & {86,47\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
        {100}    & {30}     & {68\%}                              & {42,79\%}                           & {4\%}                                    & {65\%}        & {87,99\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
        {500}    & {10}     & {83\%}                              & {79,95\%}                           & {6\%}                                    & {67\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
        {500}    & {15}     & {83\%}                              & {79,5\%}                            & {7\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
        {500}    & {30}     & {84\%}                              & {83,43\%}                           & {5\%}                                    & {68\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
        {1000}   & {10}     & {84\%}                              & {82,42\%}                           & {5\%}                                    & {67\%}        & {88,99\%}   & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
        {1000}   & {15}     & {85\%}                              & {82,42\%}                           & {5\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
        {1000}   & {30}     & {85\%}                              & {82,81\%}                           & {4\%}                                    & {59\%}        & {85,5\%}    & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
        {1500}   & {10}     & {84\%}                              & {81,42\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
        {1500}   & {15}     & \textbf{85\%}                       & \textbf{84,43\%}                    & \textbf{4\%}                             & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
        {1500}   & {30}     & {84\%}                              & {80,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
        {2000}   & {10}     & {83\%}                              & {82,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
        {2000}   & {15}     & {84\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
        {2000}   & {30}     & {85\%}                              & {82,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
        {3000}   & {10}     & {82\%}                              & {79,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
        {3000}   & {15}     & {85\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
        {3000}   & {30}     & {83\%}                              & {80,89\%}                           & {3\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
        \bottomrule
    \end{tabular}
    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
    \label{tab:results_masked-caption-generator}
    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1. It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
    % DK: might be the case. in the last language game i use only a linear layer instead of convolutions, but it doesn't help. This however might also be due to other problems. Time? (QUESTION, future work)
    % SD: I would present these for individual classes.
    % DK: I only have accuracy scores at. Time? (QUESTION)
\end{table}
